### 2015-06-01 abstract.tex
Copied files from svn

%!TEX root =  matrix-est.tex
We propose an efficient procedure for estimating the mixing time of a Markov chain.
The novelty of our approach lies in several key aspects.
Crucially,
we do not assume any a priori bounds
on the mixing time,
which
sets our approach apart from
most others that do
require some advance knowledge of this quantity.
Another
advantage of our result over
common practice is the sufficiency of a single long sampling sequence,
without access to a restart mechanism.
Perhaps most importantly, we overcome an obstacle inherent in the chicken-and-egg nature of
estimating mixing coefficients: to give confidence bounds one typically needs to know the mixing time.
We provide confidence guarantees that are fully data-dependent in the sense that they yield non-trivial
finite-sample deviation bounds without any knowledge of the mixing coefficients.
Upper and lower sample complexity bounds are provided and challenging open problems are posed.