### 2015-06-01 matrix-est.tex

%\documentclass[11pt]{article}
%\documentclass[anon,12pt]{colt2015} 
\documentclass{article}
\usepackage{nips15submit_e,times}
\input{__includes}
\renewcommand\citet\citep % since NIPS is using numbered refs

%\title{Single sample path based estimation of the mixing time in finite, reversible Markov chains}
\title{%
  Mixing time estimation in reversible Markov chains from a single
  sample path%
}

\author{
David S.~Hippocampus\thanks{ Use footnote for providing further information
about author (webpage, alternative address)---\emph{not} for acknowledging
funding agencies.} \\
Department of Computer Science\\
Cranberry-Lemon University\\
Pittsburgh, PA 15213 \\
\texttt{hippo@cs.cranberry-lemon.edu} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\AND
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
(if needed)\\
}

\begin{document}
\maketitle

\begin{abstract} 
\input{abstract}
\end{abstract}

\if0
\begin{keywords}
  Markov chain; mixing time; spectral gap;
  fully data-dependent confidence interval
\end{keywords}
\fi

\section{Introduction}\label{sec:intro}
\input{intro-nips}

\section{Notation}\label{sec:notation}
\input{notation}

\section{Setting}\label{sec:setting}
\input{setting}

%\section{Results}
%\input{results}

\section{Convergence rates}\label{sec:rates}
\input{rates}

\section{Fully empirical confidence intervals}\label{sec:empirical}
\input{empirical}

\section{Discussion}\label{sec:discussion}
\input{conclusions}

\subsubsection*{References}
{\def\section*#1{}\small \bibliography{all} \bibliographystyle{myunsrt}}

\newpage

\appendix

\section{Proof of \cref{thm:err}}\label{app:upper}
\input{upperbound-new}

\section{Proofs of lower bounds}\label{app:lower}
\input{lowerbound}

\section{Proof of \cref{thm:empirical}}\label{app:empirical}
\input{appendix-empirical}

\section{Proof of \cref{thm:combined}}\label{app:combined}
\input{appendix-combined}

\end{document}