### 2015-06-01 scratch/beta-mixing.tex

\documentclass[11pt]{article}
\usepackage{todonotes}
\newcommand{\todoc}[2][]{\todo[size=\scriptsize,color=green!10!white,#1]{Cs: #2}} % Csaba's comments
\newcommand{\todot}[2][]{\todo[size=\scriptsize,inline,color=blue!20!white,#1]{D: #2}} % Daniel's comments
\newcommand{\todor}[2][]{\todo[size=\scriptsize,color=orange!20!white,#1]{A: #2}} % Aryeh's comments

\usepackage{amsmath,amsbsy,amsfonts,amssymb,amsthm,color,dsfont}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{latexsym}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{accents}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
\usepackage{dsfont}
\usepackage[bf]{caption}
\usepackage{hyperref}
\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=false,          % non-Latin characters in AcrobatÕs bookmarks
    pdftoolbar=true,        % show AcrobatÕs toolbar?
    pdfmenubar=true,        % show AcrobatÕs menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={My title},    % title
    pdfauthor={Author},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={keyword1} {key2} {key3}, % list of keywords
    pdfnewwindow=true,      % links in new window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=blue,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}
\usepackage{amsthm}
\usepackage{times}
\usepackage{natbib}
\usepackage{nicefrac}
\usepackage{wrapfig}
\usepackage[capitalize]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\defined}{\vcentcolon =}
\newcommand{\rdefined}{=\vcentcolon}
\newcommand{\E}{\mathbb E}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\calF}{\mathcal F}
\newcommand{\sr}[1]{\stackrel{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\ind}[1]{\mathds{1}\!\!\set{#1}}
\newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand{\floor}[1]{\left \lfloor {#1} \right\rfloor}
\newcommand{\ceil}[1]{\left \lceil {#1} \right\rceil}

\def\subsubsect#1{\vspace{1ex plus 0.5ex minus 0.5ex}\noindent{\bf\boldmath{#1.}}}

\renewcommand{\P}[1]{\mathbb{P}\left\{#1\right\}}
\newcommand{\Prob}[1]{\mathbb{P}\left\{#1\right\}}
\newcommand{\EE}[1]{\mathbb{E}\left[#1\right]}

\let\temp\epsilon
\let\epsilon\varepsilon
\newcommand{\eps}{\varepsilon}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}

\newcommand{\R}{\mathds{R}}

\newcommand\wh{\ensuremath{\widehat}}
\newcommand\wt{\ensuremath{\widetilde}}
\newcommand\norm[1]{\left\| #1 \right\|}
\newcommand\tvnorm[1]{\left\| #1 \right\|_{\mathrm{TV}}}
\newcommand\parens[1]{\left( #1 \right)}
\DeclareMathOperator{\Diag}{Diag}
\DeclareMathOperator{\Sym}{Sym}
\renewcommand\t{{\ensuremath{\scriptscriptstyle{\top}}}}
\newcommand\be{\ensuremath{\mathbf{e}}}
\newcommand\tmix{\ensuremath{t_{\mathrm{mix}}}}
\newcommand\htmix{\ensuremath{\hat{t}_{\mathrm{mix}}}}
\newcommand{\od}{\bar{d}}
\newcommand{\tcouple}{\tau_{\mathrm{couple}}}
\newcommand{\ZZ}{\mathcal{Z}}
\newcommand\trel{\ensuremath{t_{\mathrm{rel}}}}
\newcommand{\ip}[1]{\langle#1\rangle}
\newcommand{\DF}{\mathcal{E}}
\newcommand\ttmix{\ensuremath{t_{\mathrm{mix},2}}}

\newcommand{\paren}[1]{\left( #1 \right)}
\newcommand{\sqprn}[1]{\left[ #1 \right]}
\newcommand{\tlprn}[1]{\left\{ #1 \right\}}
%\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\oo}[1]{\frac{1}{#1}}

\def\ddefloop#1{\ifx\ddefloop#1\else\ddef{#1}\expandafter\ddefloop\fi}
% \bbA, \bbB, ...
\def\ddef#1{\expandafter\def\csname bb#1\endcsname{\ensuremath{\mathbb{#1}}}}
\ddefloop ABCDEFGHIJKLMNOPQRSTUVWXYZ\ddefloop

\title{Statistical Estimation of Mixing Times}
\author{Daniel, Aryeh, Csaba}

\begin{document}
\maketitle

\begin{abstract}
An abstract
\end{abstract}

\section{Beta coefficients of reversible, irreducible, stationary Markov chains}

Suppose our Markov chain is reversible, irreducible and stationary (RIS),
with stationary distribution $\pi$. We would like to upper-bound its $k$th $\beta$-mixing coefficient.
The latter may be defined,
for any stationary sequence $\set{Z_t}_{t=-\infty}^{\infty}$,
 as follows:
$$ \beta(k) = \sup_n \E_B \sqprn{ \sup_A |\P{A|B}-\P{A}|},
$$
where $B$ is an event in the $\sigma$-algebra of $Z_{-\infty}^n$ and $A$ an event in $\sigma(Z_{n+k}^\infty)$.
Clearly, $\beta(k)\le\phi(k)$, where the $\phi$-mixing coefficient is defined analogously to $\beta$,
with $\sup_B$ in place of
$\E_B$  \cite{bradley05}.

For a RIS chain, we have
$$ \tmix(\eps) \le  \oo{1-\lambda_2}\ln\left( \frac{1}{\eps \pi_* }\right).$$

Inverting for $\eps$, we have
$$ \beta(k) \le \phi(k) \le \oo{\pi_*}e^{-k(1-\lambda_2)}.$$

\bibliographystyle{plainnat}
\bibliography{../all}

\end{document}