### 2015-06-01 scratch/synthesis.tex

\documentclass[11pt]{article}
\usepackage{todonotes}
\newcommand{\todoc}[2][]{\todo[size=\scriptsize,color=green!10!white,#1]{Cs: #2}} % Csaba's comments
\newcommand{\todot}[2][]{\todo[size=\scriptsize,inline,color=blue!20!white,#1]{D: #2}} % Daniel's comments
\newcommand{\todor}[2][]{\todo[size=\scriptsize,color=orange!20!white,#1]{A: #2}} % Aryeh's comments

\usepackage{amsmath,amsbsy,amsfonts,amssymb,amsthm,color,dsfont}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{latexsym}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{accents}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
\usepackage{dsfont}
\usepackage[bf]{caption}
\usepackage{hyperref}
\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=false,          % non-Latin characters in AcrobatÕs bookmarks
    pdftoolbar=true,        % show AcrobatÕs toolbar?
    pdfmenubar=true,        % show AcrobatÕs menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={My title},    % title
    pdfauthor={Author},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={keyword1} {key2} {key3}, % list of keywords
    pdfnewwindow=true,      % links in new window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=blue,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}
\usepackage{amsthm}
\usepackage{times}
\usepackage{natbib}
\usepackage{nicefrac}
\usepackage{wrapfig}
\usepackage[capitalize]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\defined}{\vcentcolon =}
\newcommand{\rdefined}{=\vcentcolon}
%\newcommand{\E}{\mathbb E}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\calF}{\mathcal F}
\newcommand{\calR}{\mathcal R}
\newcommand{\sr}[1]{\stackrel{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\ind}[1]{\mathds{1}\!\!\set{#1}}
\newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand{\floor}[1]{\left \lfloor {#1} \right\rfloor}
\newcommand{\ceil}[1]{\left \lceil {#1} \right\rceil}

\def\subsubsect#1{\vspace{1ex plus 0.5ex minus 0.5ex}\noindent{\bf\boldmath{#1.}}}

\renewcommand{\P}[1]{\mathbb{P}\left\{#1\right\}}
\newcommand{\Prob}[1]{\mathbb{P}\left\{#1\right\}}
\newcommand{\EE}[1]{\mathbb{E}\left[#1\right]}

\let\temp\epsilon
\let\epsilon\varepsilon
\newcommand{\eps}{\varepsilon}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}

\newcommand{\R}{\mathds{R}}

\newcommand\wh{\ensuremath{\widehat}}
\newcommand\wt{\ensuremath{\widetilde}}
\newcommand\norm[1]{\left\| #1 \right\|}
\newcommand\tvnorm[1]{\left\| #1 \right\|_{\mathrm{TV}}}
\newcommand\parens[1]{\left( #1 \right)}
\DeclareMathOperator{\Diag}{Diag}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator*{\E}{\text{\bf E}}
\renewcommand\t{{\ensuremath{\scriptscriptstyle{\top}}}}
\newcommand\be{\ensuremath{\mathbf{e}}}
\newcommand\tmix{\ensuremath{t_{\mathrm{mix}}}}
\newcommand\htmix{\ensuremath{\hat{t}_{\mathrm{mix}}}}
\newcommand{\od}{\bar{d}}
\newcommand{\tcouple}{\tau_{\mathrm{couple}}}
\newcommand{\ZZ}{\mathcal{Z}}
\newcommand\trel{\ensuremath{t_{\mathrm{rel}}}}
\newcommand{\ip}[1]{\langle#1\rangle}
\newcommand{\DF}{\mathcal{E}}
\newcommand\ttmix{\ensuremath{t_{\mathrm{mix},2}}}

\newcommand{\paren}[1]{\left( #1 \right)}
\newcommand{\sqprn}[1]{\left[ #1 \right]}
\newcommand{\tlprn}[1]{\left\{ #1 \right\}}
%\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\oo}[1]{\frac{1}{#1}}
\newcommand{\nrm}[1]{\left\Vert #1 \right\Vert}
\newcommand{\tnrm}[1]{\Vert #1 \Vert}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\trn}{^\intercal} %operator transpose

\newcommand{\beq}{\begin{eqnarray*}}
\newcommand{\eeq}{\end{eqnarray*}}
\newcommand{\beqn}{\begin{eqnarray}}
\newcommand{\eeqn}{\end{eqnarray}}

\def\ddefloop#1{\ifx\ddefloop#1\else\ddef{#1}\expandafter\ddefloop\fi}
% \bbA, \bbB, ...
\def\ddef#1{\expandafter\def\csname bb#1\endcsname{\ensuremath{\mathbb{#1}}}}
\ddefloop ABCDEFGHIJKLMNOPQRSTUVWXYZ\ddefloop

\title{Synthesis..}
%\author{Daniel, Aryeh, Csaba}

\begin{document}
\maketitle

%\begin{abstract}
%An abstract
%\end{abstract}
\newcommand{\XX}{\mathbb{X}}
\newcommand{\one}[1]{\mathbb{I}\left\{#1\right\}}
\newcommand{\N}{\mathbb{N}}
Let  $X_1,\ldots,X_n$ be a  Markov chain with transition matrix $T$
over the state space $\XX = \{1,\ldots,d \}$.
We assume that the chain is reversible, irreducible, and stationary (RIS).
\todoc[inline]{Fix this: We don't want stationarity.}
Our goal is to derive a hight probability bound on the deviation 
$\bar{X}_n - \EE{f(X_1)} = \frac1n \sum_{t=1}^n f(X_t) - \EE{f(X_1)}$
(here, $f$ is some $[0,1]$-valued function).
We know that $\bar{X}_n - \EE{f(X_1)}$ can be bounded in terms of the $\beta$-mixing coefficients of $(X_i)$,
following the ``blocking'' method of \cite{MR1258867}.
In particular, for any $\mu,a$ such that $n = 2\mu a$ and $\delta>2(\mu-1)\beta(a)$,
with probability at least $1-\delta$,
\begin{align}
\label{eq:xbound}
|\bar{X}_n - \EE{f(X_1)}|
\le c_1\sqrt{\frac{1}{\mu}} + c_2\sqrt{\frac{\ln(2/\delta')}{2\mu}}\,,
\end{align}
where $\delta'=\delta-2(\mu-1)\beta(a)$ (see mohri-notes.pdf).
Now note that if $\beta(k)$ denotes the $k$th $\beta$-mixing coefficient of $(X_i)$ then 
\begin{align}
\label{eq:betabound}
 \beta(k) \le \oo{\pi_*}e^{-k \gamma^*}
\end{align}
where  $\gamma^* = 1-\lambda^*$ is the spectral gap ($\lambda^*$ is the second largest eigenvalue of $T$)
and  $\pi^*  = \min_i \pi_i>0$ is the critical probability underlying the chain
(see the $\beta$-mixing note).
For a given value of $\pi^*,\gamma^*$, $\delta$, let $B_n(\pi^*,\gamma^*,\delta)$ be the optimal value of the optimization problem
\begin{align*}
\lefteqn{ \min_{\mu,a} \min\left(1, c_1\sqrt{\frac{1}{\mu}} + c_2\sqrt{\frac{\ln(2/\delta')}{2\mu}}\right) } \\
& \qquad \text{s.t.} \quad n = 2\mu a, \quad \delta>2(\mu-1)\oo{\pi_*}e^{-a \gamma^*},
\quad \delta' = \delta-2(\mu-1)\oo{\pi_*}e^{-a \gamma^*}, \quad \mu,a \in \N\,.
\end{align*}

Thus, we need estimates of $\gamma^*$ and $\pi^*$.
To estimate them we may want to estimate the stationary probabilities and the transition matrix $T$.
Let $\hat \pi$ denote the estimate of the former, while let $\hat T$ be the estimate of the latter.
In particular, $\hat \pi_i$ is taken to be the empirical frequency of the visits of state $i$ in $(X_t)$,
while we let $\hat T_{i,j}$ the frequency when 
$X_{t+1}=j$ happens given that $X_t = i$ ($t=1,\dots,n-1$). 
When $X_t = i$ never happens, we take $\hat T_{i,\cdot}$ to be the uniform distribution.
Since $\Prob{\sum_{t=1}^n \one{X_t = i}=0}=\Prob{X_1\ne i, \dots, X_n\ne i}>0$ for each $i$, $\hat T_{ij}$ is a biased estimate of $T_{ij}$, although the bias decreases exponentially fast.

For $x\in \R^d$, let $D_x$ be the diagonal matrix with $x$ in its diagonal. 
Define $L = D_\pi^{-1/2} T  D_\pi^{1/2}$, $\hat{L} = D_{\hat{\pi}}^{-1/2} \hat{T} D_{\hat{\pi}}^{1/2}$, 
By Weyl's inequality (see notes.pdf),
\begin{align}
\label{eq:gammabound}
|\hat{\gamma}^* - \gamma^* | = |\hat{\lambda}^* - \lambda^* | \le \tnrm{\hat L - L }_2
\end{align}
and so it remains to bound this latter quantity.
Define $E_T = D_{\pi}^{-1/2} (\hat T - T ) D_{\pi}^{1/2}$
and $E_\pi = D_{\hat{\pi}}^{-1/2} D_\pi^{1/2} - I$.
We have (see notes.pdf),
\begin{align}
\label{eq:lbound}
\tnrm{\hat L - L}_2 \le \norm{E_T}_2 + \norm{E_{\pi}}_2 (\norm{E_T}_2+1) (\norm{E_\pi}_2+2)\,.
\end{align}
Now, $\norm{E_T}_2 = \tnrm{\hat T - T}_2$ (because the eigenvalues of $E_T$ are the same as that of $\hat T - T$).
Further, for any $\mu,a$ such that $n=2\mu a$
and $\delta>2(\mu-1)\beta(a)$,
with probability at least $1-\delta$,
\begin{align}
\label{eq:tbound}
\tnrm{T-\hat T}_2
\le 25\sqrt{\frac{d}{\mu}} + 2\sqrt{\frac{\ln(2/\delta')}{2\mu}}\,,
\end{align}
where $\delta'=\delta-2(\mu-1)\beta(a)$ (see mohri-notes.pdf). \todoc{This does not take into account the bias.}
\todoc[inline]{Next problem: For a fixed $n$, this will limit the range of $\delta$.
For any $n$, there will be a range of $\delta$ in the neighbourhood of $0$ for which the algorithm
will need to return ``I don't know''.
}
Similarly, for any $\mu',a'$ such that $n=2\mu' a'$
and $\delta>2(\mu'-1)\beta(a')$,
with probability at least $1-\delta$,
\begin{align}
\label{eq:pibound}
|\hat\pi_i - \pi_i | \le C \left( \sqrt{\frac{d}{\mu'}} + \sqrt{\frac{\ln(2/\delta'')}{2\mu'}}\right)\,,
\end{align}
where $\delta''=\delta-2(\mu'-1)\beta(a')$.

By combining~\eqref{eq:betabound},
\eqref{eq:gammabound}, \eqref{eq:lbound}, \eqref{eq:tbound} and~\eqref{eq:pibound}, we can derive 
that for any $0\le \delta \le 1$,
 $\mu,a,\mu',a'$  such that $n = 2\mu a$ and
 $\frac{2}{\pi^*} \max( (\mu-1)e^{-a\gamma^*}, (\mu'-1)e^{-a'\gamma^*} )<\delta$, 
with probability $1-\delta$,
\begin{align*}
|\hat{\gamma}^* - \gamma^* | &\le u( \gamma^*, \pi^*, \hat{\gamma}, \hat\pi, \hat T, n, \mu, a, \mu', a', \delta )\,,\\
|\hat{\pi}^* - \pi^* | &\le v( \gamma^*, \pi^*, \hat{\gamma}, \hat\pi, \hat T, n, \mu, a, \mu', a', \delta )\,,
\end{align*}
for some functions $u$ and $v$.
Reordering (and relaxing) the  inequalities we get
\begin{align*}
\gamma^* &\ge \hat\gamma^* - u( \gamma^*, \pi^*, \hat{\gamma}, \hat\pi, \hat T, n, \mu, a, \mu', a', \delta )\,,\\
 \pi^*  &\ge \hat{\pi}^* - v( \gamma^*, \pi^*, \hat{\gamma}, \hat\pi, \hat T, n, \mu, a, \mu', a', \delta )\,.
\end{align*}
Now, let $\hat{B}_n(\delta)$ be the optimal value of the optimization problem 
\begin{align*}
\lefteqn{\max B_n(g,p,\delta)\qquad \text{s.t.} }\\
&\qquad  g \ge \hat\gamma^* - u( g, p, \hat{\gamma}, \hat\pi, \hat T, n, \mu, a, \mu', a', \delta )\,,\\
&\qquad p  \ge \hat{\pi}^* - v( g, p, \hat{\gamma}, \hat\pi, \hat T, n, \mu, a, \mu', a', \delta )\,,\\
&\qquad n = 2\mu a, \quad n = 2\mu' a'\,,\\
&\qquad \frac{2}{p} \max( (\mu-1)e^{-ag}, (\mu'-1)e^{-a'g} )<\delta\,,\\
&\qquad g,p >0, \mu,a,\mu',a'\in \N\,.
\end{align*}
Then, w.p. $1-\delta$, it holds that
\[
|\bar{X}_n - \EE{f(X_1)}| \le \hat{B}_n(\delta)\,.
\]

Next goals: Analytic upper bound on $\EE{ \hat{B}_n(\delta)}$.
Matching lower bound.

\bibliographystyle{plainnat}
\bibliography{all}

\end{document}