### 2015-06-01 setting.tex
adding and resolving todos, as per our discussion.
bib file improvements
generalized inverse clean up, both in appendix and in main text
ideas for future work in conclusion
empirical results: comment on O(.) notation
minor tweak to intro

%!TEX root =  matrix-est.tex
\label{sec:setting}
Let $\vP \in (\Delta^{d-1})^d \subset [0,1]^{d \times d}$ be a $d
\times d$ row-stochastic matrix for an ergodic (i.e., irreducible and
aperiodic) Markov chain.
This implies there is a unique stationary distribution $\vpi \in
\Delta^{d-1}$ with $\pi_i > 0$ for all $i \in [d]$~\citep[Corollary
1.17]{LePeWi08}.
We also assume that $\vP$ is \emph{reversible} (with respect to
$\vpi$):
\begin{align}
\label{eq:reversibility}
  \pi_i P_{i,j} = \pi_j P_{j,i} ,
  \quad i,j \in [d] .
\end{align}
The minimum stationary probability is denoted by $\pimin := \min_{i
\in [d]} \pi_i$.

Define the matrices
\begin{align*}
\vM := \Diag(\vpi) \vP \quad \text{and} \quad
\vL := \Diag(\vpi)^{-1/2} \vM \Diag(\vpi)^{-1/2}\,.
\end{align*}
The $(i,j)$th entry of the matrix $M_{i,j}$ contains the \emph{doublet
probabilities} associated with $\vP$: $M_{i,j} = \pi_i P_{i,j}$ is the
probability of seeing state $i$ followed by state $j$ when the chain
is started from its stationary distribution.
The matrix $\vM$ is symmetric on account of the reversibility of
$\vP$, and hence it follows that $\vL$ is also symmetric.
As we will see later, we will strongly exploit this later.  \todoc{Say more? Maybe not..}
%will be advantageous due to the existence of stronger perturbation results
%for symmetric matrices.
Further, $\vL = \Diag(\vpi)^{1/2} \vP \Diag(\vpi)^{-1/2}$, hence $\vL$
and $\vP$ are similar and thus their eigenvalue systems are identical.
Ergodicity and reversibility imply that the eigenvalues of $\vL$ are
contained in the interval $\opencloseint{-1,1}$, and that $1$ is an
eigenvalue of $\vL$ with multiplicity $1$~\citep[Lemmas 12.1 and
12.2]{LePeWi08}.
Denote and order the eigenvalues of $\vL$ as
\[
  1 = \lambda_1 > \lambda_2 \geq \dotsb \geq \lambda_d > -1 .
\]
Let $\slem := \max\{ \lambda_2,\, |\lambda_d| \}$, and define the
(absolute) spectral gap to be $\gap := 1-\slem$, which is strictly positive.
\todoc{Do our results generalize to the spectral gap? Where do we need that we talk about the absolute spectral gap?
What is the difference between the spectral and the absolute spectral gap? Say, in their relationship to $\tmix$?}
%Throughout the paper, both $\pimin$ and $\gap$ are assumed
%to be positive.

Let $(X_t)_{t\in\bbN}$ be a Markov chain whose transitions are governed by
$\vP$.
For each $t \in \bbN$, let $\vpi^{(t)} \in \Delta^{d-1}$ denote the
marginal distribution of $X_t$, so
\[
  \vpi^{(t+1)} = \vpi^{(t)} \vP ,
  \quad t \in \bbN .
\]
Note that the initial distribution $\vpi^{(1)}$ is arbitrary,
and need not be the stationary distribution $\vpi$.

The goal is to estimate $\pimin$ and $\gap$ from the length $n$ sample
path $(X_t)_{t \in [n]}$, and also to construct fully empirical
confidence intervals that $\pimin$ and $\gap$ with high probability;
in particular, the construction of the intervals should not depend on
any unobservable quantities, including $\pimin$ and $\gap$ themselves.
It is well-known that the \emph{mixing time} of the Markov chain $\tmix$ (defined in the introduction)
\hide{,
\[
  \tmix
  :=
  \min\Braces{
    t \in \bbN :
    \sup_{\vq \in \Delta^{d-1}}
    \Norm{
      \vq \vP^t - \vpi
    }_{\tv} \leq 1/4
  }
  ,
\]}
can be bounded in terms of $\pimin$ and $\gap$~\citep[Theorems~12.3
and~12.4]{LePeWi08}:
\[
  \Parens{\frac1{\gap} - 1} \ln2
  \ \leq \ \tmix
  \ \leq \ \frac1{\gap} \log \frac4{\pimin}
  .
\]
Moreover, convergence rates for empirical processes on Markov chain
sequences are also often given in terms of mixing coefficients that
can ultimately be bounded in terms of $\pimin$ and
$\gap$, as we will show in the proof of our first result.\todoc{I removed the citations (saving space), referring to the proof.}
%~\citep[e.g.,][]{MR1921877,MoRo08,MoRo09}. 
Therefore, valid confidence intervals for $\pimin$ and $\gap$ can be
used to make these rates fully observable.

%Therefore, confidence intervals for $\pimin$ and $\gap$ can be used to
%make these rates data-dependent and hence usable in procedures such
%as Structural Risk Minimization.
%\todoc{citation for SRM, and other applications?}
%
%Additionally, we aim to provide
%asymptotic convergence rates
%as well as data-dependent confidence intervals;
%the latter should not depend on any unknown quantities
%(such as $\pimin$ and $\gap$).
%
%\todoc[inline]{What is the data? What is the problem?}