### 2015-06-02 empirical.tex

%!TEX root =  matrix-est.tex
In this section, we address the shortcoming of \cref{thm:err} and give
fully empirical confidence intervals for the stationary probabilities
and the spectral gap $\gap$.
The main idea is to use the Markov property to eliminate the
dependence of the confidence intervals on the unknown quantities
(including $\pimin$ and $\gap$).
Specifically, we estimate the transition probabilities from the sample
path using simple frequency estimates: as a consequence of the Markov
property, for each state, the frequency estimates converge at a rate
that depends only on the number of visits to the state, and in
particular the rate (given the visit count of the state) is
independent of the mixing time of the chain.

As discussed in \cref{sec:rates}, it is possible to form a confidence
interval for $\gap$ based on the eigenvalues of an estimated
transition probability matrix by appealing to the
Ostrowski-Elsner theorem.
However, as explained earlier, this would lead to a slow
$O(n^{-1/(2d)})$ rate.
We avoid this slow rate by using an estimate of the symmetric matrix
$\vL$, so that we can use a stronger perturbation result (namely Weyl's
inequality, as in the proof of \cref{thm:err}) available for symmetric matrices.

To form an estimate of $\vL$ based on an estimate of the transition
probabilities, one possibility is to estimate $\vpi$ using a
frequency-based estimate for $\vpi$ as was done in \cref{sec:rates},
and appeal to the relation $\vL = \Diag(\vpi)^{1/2} \vP
\Diag(\vpi)^{-1/2}$ to form a plug-in estimate.
However, as noted in \cref{sec:rates-upper}, confidence intervals for
the entries of $\vpi$ formed this way may depend on the mixing time.
Indeed, such an estimate of $\vpi$ does not exploit the Markov
property.

We adopt a different strategy for estimating $\vpi$, which leads to
our construction of empirical confidence intervals, detailed in
\cref{alg:empest}.
We form the matrix $\wh\vP$ using smoothed frequency estimates of
$\vP$ (Step~\ref{step:P}), then compute the so-called group inverse
$\giAh$ of $\wh\vA = I - \wh\vP$ (Step~\ref{step:gi}), followed by
finding the unique stationary distribution $\hat\vpi$ of $\wh\vP$
(Step~\ref{step:pi}), this way decoupling the bound on the accuracy of $\hat\vpi$
from the mixing time.
The group inverse $\giAh$ of $\wh\vA$ is uniquely defined;
 and if
$\wh\vP$ defines an ergodic chain (which is the case here due to the
use of the smoothed estimates), $\giAh$ can be computed at the cost of
inverting an $(d-1)\times (d-1)$ matrix~\citep[Theorem
5.2]{meyer1975role}.%
\footnote{
\label{ftnote:group-inverse}
The group inverse of a square matrix $\vA$, a special case of the {\em Drazin inverse},
is the unique matrix $\vA^\#$ satisfying
$ \vA\vA^\#\vA = \vA$,
$\vA^\#\vA\vA^\# = \vA^\#$ and
$\vA^\#\vA = \vA\vA^\#$.
%It is efficiently computable,
%and for $\vA=\vI-\vP$, where $\vP$ is ergodic with stationary distribution $\vpi$,
%has the
%following simple
%form: $\vA^\#=(\vI-\vP+\v1\vpi^\t)^{-1}-\v1\vpi^\t$
%\citep{seneta1993sensitivity,meyer1975role}.
}
Further, once given $\giAh$, the unique stationary distribution
$\hat\vpi$ of $\wh\vP$ can be read out from the last row of
$\giAh$~\citep[Theorem 5.3]{meyer1975role}.
The group inverse will also be used later to compute the sensitivity
of $\vpi$ based on a result of Cho and Meyer \cite{cho2001comparison}.
Based on $\hat\vpi$ and $\wh\vP$, we construct the plug-in estimate
$\wh\vL$ of $\vL$, and use the eigenvalues of its symmetrization to
form the estimate $\hatgap$ of the spectral gap (Steps~\ref{step:eig}
and~\ref{step:gap}).
In the remaining steps, 
we use perturbation analyses to relate $\hat\vpi$ and $\vpi$, viewing
$\vP$ as the perturbation of $\wh\vP$; and also to relate $\hatgap$
and $\gap$, viewing $\vL$ as a perturbation of $\Sym(\wh\vL)$.
Both analyses give error bounds entirely in terms of observable
quantities, tracing back to empirical error bounds for the smoothed
frequency estimates of $\vP$.
Note that the most expensive step in the algorithm is the computation of the group inverse,
which, as noted reduces to matrix inversion.
Thus, with a standard implementation of matrix inversion, the algorithm's complexity becomes 
$O( n + d^3)$, while its space complexity is $O(d^2)$.
\todoc{Implement and test?}

\begin{algorithm}[t]
\caption{Empirical confidence intervals}
\label{alg:empest}
\begin{algorithmic}[1]
  \renewcommand\algorithmicrequire{\textbf{Input}:}
  \REQUIRE
    Sample path $(X_1,X_2,\dots,X_n)$,
    confidence parameter $\delta \in (0,1)$.

  \STATE Compute state visit counts and smoothed transition
  probability estimates:
  \begin{equation}
    \begin{aligned}
      N_i & :=
      \Abs{
        \Braces{
          t \in [n-1] : X_t = i
        }
      }
      , \quad i \in [d] ; \\
      N_{i,j} & :=
      \Abs{
        \Braces{
          t \in [n-1] : (X_t,X_{t+1}) = (i,j)
        }
      }
      , \quad (i,j) \in [d]^2 ; \\
      \wh{P}_{i,j}
      & :=
      \frac{N_{i,j} + 1/d}{N_i + 1}
      , \quad (i,j) \in [d]^2 .
    \end{aligned}
    %\label{eq:frequencies}
    \notag
  \end{equation}
  \label{step:P}

  \STATE Let $\giAh$ be the group inverse of $\wh\vA := \vI -
  \wh\vP$.
  \label{step:gi}

  \STATE Let $\hat\vpi \in \Delta^{d-1}$ be the unique stationary
  distribution for $\wh\vP$.
  \label{step:pi}

  \STATE Compute eigenvalues $\hat\lambda_1 {\geq} \hat\lambda_2
  {\geq} \dotsb {\geq} \hat\lambda_d$ of $\Sym(\wh\vL)$, where $\wh\vL
  := \Diag(\hat\vpi)^{1/2} \wh\vP \Diag(\hat\vpi)^{-1/2}$.
  \label{step:eig}

  \STATE Spectral gap estimate:
  \[ \hatgap := 1 - \max\braces{ \hat\lambda_2,\, |\hat\lambda_d| } . \]
  \label{step:gap}

  \STATE Empirical bounds for $|\wh{P}_{i,j}-P_{i,j}|$ for each $(i,j)
  \in [d]^2$:
  \begin{equation}
    \wh{B}_{i,j}
    :=
    \sqrt{\frac{9\wh{P}_{i,j}(1-\wh{P}_{i,j})\ln(2d/\delta)}{2N_i}}
    + \frac{4\ln(2d/\delta)+1.5\abs{(1/d)-\wh{P}_{i,j}}}{N_i}
    .
%    \label{eq:P-obs-bound}
    \notag
  \end{equation}
  \label{step:P-bound}

%$\qquad$ $\wh\Delta_{i,j}(\delta) \gets \sqrt{\frac{9\wh{P}_{i,j}(1-\wh{P}_{i,j})\ln(2d/\delta)}{2N_i}}
%  + \frac{4\ln(2d/\delta)+1.5\abs{\alpha-d\alpha\wh{P}_{i,j}}}{N_i}$%,  $(i,j) \in [d]^2$.
%\STATE   $\qquad$ $\wh\Delta_{i,j}(\delta) \gets \min( \wh\Delta_{i,j}(\delta) , \max(\wh{P}_{i,j}, 1-\wh{P}_{i,j}))$,   $(i,j) \in [d]^2$.

%  \STATE Empirical bound for $\max_{i,j}|\wh{P}_{i,j}-P_{i,j}|$:
%  \quad $B(\delta) := \max_{i,j} B_{i,j}(\delta)$.
  %$\wh\Delta_{\vP}(\delta) \gets \sqrt{\sum_{i,j} \wh\Delta_{i,j}^2(\delta)}$,
%  $\wh\Delta^*_{\vP}(\delta) \gets \max_{i,j} \wh\Delta_{i,j}(\delta)$.

  \STATE Relative sensitivity of $\vpi$:
  \begin{equation}
    \hat\kappa :=
    \frac12
    \max
    \Braces{
      \wh{A}_{j,j}^\# - \min\Braces{ \wh{A}_{i,j}^\# : i \in [d] }
      : j \in [d]
    } 
    .
%    \label{eq:hatkappa}
    \notag
  \end{equation}
  \label{step:sens}

  \STATE Empirical bounds for $\max_{i \in [d]} |\hat{\pi}_i -
  \pi_i|$ and
  $\max\bigcup_{i\in[d]}
  \braces{
    \abs{\sqrt{\pi_i/\hat\pi_i}-1},\,
    \abs{\sqrt{\hat\pi_i/\pi_i}-1}
  }$:
  \begin{equation}
    \hat{b} := \hat\kappa \max\Braces{
      \wh{B}_{i,j}
      : (i,j) \in [d]^2
    }
    , \qquad
    \hat\rho := \frac12 \max \bigcup_{i\in[d]}
    \Braces{
      \frac{\hat{b}}{\hat\pi_i},\,
      \frac{\hat{b}}{\brackets{\hat\pi_i-\hat{b}}_+}
    }
    .
%    \label{eq:pi-obs-bound}
    \notag
  \end{equation}
  \label{step:pi-bound}

  \STATE Empirical bounds for $\abs{\hatgap-\gap}$:
  \begin{equation}
    \hat{w} := 2\hat\rho + \hat\rho^2
    + (1+2\hat\rho+\hat\rho^2)
    \Biggl(
      \sum_{(i,j)\in[d]^2} \frac{\hat\pi_i}{\hat\pi_j} \hat{B}_{i,j}^2
    \Biggr)^{1/2} .
%    \label{eq:gap-obs-bound}
    \notag
  \end{equation}
  \label{step:gap-bound}

%  \STATE Empirical bounds for $\hat{\pi}_i$:
%  \begin{align*}
%    \rho_i(\delta)
%    & := \frac{\wh\kappa B(\delta)}{\hat\pi_i}
%    , \quad i \in [d] ;
%    &
%    \rho_i'(\delta)
%    & := \frac{\wh\kappa B(\delta)}{[\hat{\pi}_i-\wh\kappa B(\delta)]_+}
%    , \quad i \in [d] .
%  \end{align*}

%\STATE $\rho(\delta) \gets \frac12 \max_{i\in \iset{d}} \rho_i(\delta)$; $\rho'(\delta) \gets \frac12 \max_{i\in \iset{d}} \rho_i'(\delta)$.
%\STATE $\hat{\rho}(\delta) \gets 2\max(\rho(\delta),\rho'(\delta)) + \max(\rho(\delta),\rho'(\delta))^2$.
%%$\hat{\rho}(\delta) \gets \rho(\delta) + \rho'(\delta) + \rho'(\delta)\rho(\delta)$. % THIS WAS BUGGY
%\STATE Half-width of interval: $w \gets 
%\hat{\rho}(\delta)+ \parens{ 1+ \hat{\rho}(\delta) } \, \sqrt{ \sum_{i,j} \frac{\hat{\pi}_i}{\hat{\pi}_j} \wh{\Delta}^2_{i,j}(\delta)}
%$.
%\STATE Confidence interval for $\gap$: $U_n(\delta) \gets [\hatgap-w,\hatgap+w]\cap [0,1]$.
\end{algorithmic}
\end{algorithm}

The main result of this section is the following result:
\begin{theorem}
  \label{thm:empirical}
  Suppose \cref{alg:empest} is given as input a sample path of length
  $n$ from an ergodic and reversible Markov chain and confidence
  parameter $\delta \in (0,1)$.
  Let $\gap>0$ denote the spectral gap, $\vpi$ the unique stationary
  distribution, and $\pimin>0$ the minimum stationary probability.
  Then, on an event of probability at least $1-\delta$,
  \[
    \pi_i \in [\hat\pi_i-\hat{b},\hat\pi_i+\hat{b}]
    \quad \text{for all $i \in [d]$} ,
    \qquad\text{and}\qquad
    \gap \in [\hatgap-\hat{w},\hatgap+\hat{w}]
    .
  \]
  Moreover, $\hat{b}$ and $\hat{w}$ almost surely satisfy (as $n \to
  \infty$)
  \[
    \hat{b}
    =
    O\Parens{
      \max_{(i,j) \in [d]^2}
      \kappa
      \sqrt{\frac{P_{i,j}\log(d/\delta)}{\pi_i n}}
    }
    ,
    \quad
    \hat{w}
    =
     O\Parens{ 
     \frac{\kappa}{\pimin} \sqrt{\frac{\ln(d/\delta)}{\pimin n}} + 
      \sqrt{ \frac{d\ln(d/\delta)}{ \pimin n}}
    }
    ,
  \]
  where $\kappa$ is analogous to the $\hat{\kappa}$ computed
  by~\cref{alg:empest} with $\giAh$ replaced by the group inverse
  $\giA$ of $\vA := \vI - \vP$.\footnote{%
    In \cref{thm:empirical,thm:combined}, our use of big-$O$ notation
    is as follows.
    For a random sequence $(Y_n)_n$ and a (non-random) positive
    sequence $(a_{\theta,n})_n$ parameterized by $\theta$, we say
    ``$Y_n = O(a_{\theta,n})$ holds almost surely as $n\to\infty$'' if
    there is some universal constant $C>0$ such that for all $\theta$,
    $\limsup_{n\to\infty} Y_n/a_{\theta,n} \leq C$ holds almost
    surely.%
  }%
%  \todoc{
%    Precisely,
%    what we mean by the $O(\cdot)$ notation is as follows:
%  for a random sequence of nonnegative numbers $(X_n)_n$
%  and another sequence of nonrandom positive numbers $(a_n^\eps)_n$ 
%  indexed by $\eps$ we
%  say that $X_n = O(a_n^\eps)$ holds almost surely as $n\to \infty$
%  if for some universal constant $C>0$, 
%  $\limsup_{n\to \infty} X_n/a_n^\eps \le C$ holds almost surely, regardless of
%  the choice of $\eps$. I think this should be made clear.
%  }
\end{theorem}
The proof of \cref{thm:empirical} is given in \cref{app:empirical}.
As mentioned above, the obstacle encountered in \cref{thm:err} is
avoided by exploiting the Markov property.
We establish fully observable upper and lower bounds on the entries of
$\vP$ that converge at a $\sqrt{n}$ rate using standard martingale
tail inequalities; this justifies the validity of the bounds from
Step~\ref{step:P-bound}.
Properties of the group inverse~\citep{meyer1975role,cho2001comparison} and eigenvalue
perturbation theory~\citep{stewart1990matrix} are used to validate the
empirical bounds on $\pi_i$ and $\gap$ developed in the remaining
steps of the algorithm.

The first part of \cref{thm:empirical} provides valid empirical
confidence intervals for each $\pi_i$ and for $\gap$, which are
simultaneously valid at confidence level $\delta$.
The second part of \cref{thm:empirical} shows that the width of the
intervals decrease as the sequence length increases.
It is known that $\kappa \le d/\gap$ (see \cref{sec:asymptotic}),
which gives
\[
  \hat{b}
  =
  O\Parens{
    \max_{(i,j) \in [d]^2}
    \frac{d}{\gap}
    \sqrt{\frac{P_{i,j}\log(d/\delta)}{\pi_i n}}
  }
  , \quad
  \hat{w}
  =
  O\Parens{ 
    \frac{d}{\pimin\gap} \sqrt{\frac{\log(d/\delta)}{\pimin n}}
  }
  .
\] 
The widths of these intervals are somewhat larger than the corresponding
(non-empirical) intervals from \cref{thm:err}.
\todoc{This needs to be updated once we finalize the results. By how much are these larger?}

An easy way to improve the asymptotic width is to combine
\cref{thm:err,thm:empirical}.
Specifically, determine lower bounds on $\pimin$ and $\gap$ using
\cref{alg:empest},
\[
  \pimin \geq \min_{i \in [d]} \brackets{ \hat\pi_i - \hat{b} }_+
  , \quad
  \gap \geq \brackets{ \hatgap - \hat{w} }_+
  ;
\]
then plug-in these lower bounds for $\pimin$ and $\gap$ in the
deviation bounds in \cref{eq:gapbound} from \cref{thm:err}.
This yields a new interval centered around the estimate of $\gap$ from
\cref{thm:err}, and it no longer depends on unknown quantities.
The interval is a valid $1-2\delta$ probability confidence interval
for $\gap$, and for sufficiently large $n$, the width shrinks at the
rate given in \cref{eq:gapbound}.
We can similarly construct an empirical confidence interval for
$\pimin$ using \cref{eq:piminbound}, which is valid on the same
$1-2\delta$ probability event.\footnote{%
  For the $\pimin$ interval, we only plug-in lower bounds on $\pimin$
  and $\gap$ only where these quantities appear as $1/\pimin$ and
  $1/\gap$ in \cref{eq:piminbound}.
  It is then possible to ``solve'' for observable bounds on $\pimin$.
  See \cref{app:combined} for details.%
}
This is summarized in the following \lcnamecref{thm:combined}, which
we prove in \cref{app:combined}.
\begin{theorem}
  \label{thm:combined}
  The following holds under the same conditions as
  \cref{thm:empirical}.
  For any $\delta \in (0,1)$, the confidence intervals $\wh{U}$ and
  $\wh{V}$ described above for $\pimin$ and $\gap$, respectively,
  satisfy $\pimin \in \wh{U}$ and $\gap \in \wh{V}$ with probability
  at least $1-2\delta$.
  Furthermore, the widths of these intervals almost surely satisfy
  (as $n \to \infty$) \todoc{Do we also care about $\pimin\to 0$, $\gap\to 0$? Do we need then other terms for both? Replicating the form of bounds from \cref{thm:err}?}
  \todod{I don't understand the first question. For the second, we
  could do this, but it would take some non-trivial work\ldots}
  \[
    |\wh{U}|
    =
    O\Parens{
      \sqrt{\frac{\pimin\log\frac{d}{\pimin\delta}}{\gap n}}
    }
    ,
    \quad
    |\wh{V}|
    =
    O\Parens{
      \sqrt{\frac{\log\frac{d}{\delta}\cdot\log(n)}{\pimin\gap n}}
    }
    .
  \]
\end{theorem}

%The rest of the section is devoted to the proof of \cref{thm:empirical}.

%In what follows, we will (re-)define  $\wh{\Delta}_{i,j}(\delta)$ to be the RHS of \eqref{eq:P-obs-bound} when $N_i>0$,
% \todoc{Abbrev RHS should be introduced.}
%while we let $\wh{\Delta}_{i,j}(\delta)=\max(\wh{P}_{i,j},1-\wh{P}_{i,j})$ otherwise.
%%Clearly, $\wt{\Delta}_{i,j}(\delta) \ge \wh{\Delta}_{i,j}(\delta)$.
%%In what follows, we calculate with the tighter $\wh{\Delta}_{i,j}(\delta)$. 
%%In fact, the algorithm was presented for $\wt{\Delta}_{i,j}(\delta)$ for easier comprehension only.
%