### 2015-06-02 abstract.tex
.

%!TEX root =  matrix-est.tex
This article provides an efficient procedure for estimating the mixing
time of a reversible ergodic Markov chain.
The procedure computes the estimate from a single finite-length sample
path from the Markov chain, and does not require any \emph{a priori}
bounds on the mixing time itself.
This stands in contrast to previous approaches, which either require a
reset mechanism or require additional prior knowledge.
Most notably, the procedure produces a fully empirical confidence
interval (i.e., it depends only on the sample path and a given
confidence parameter $\delta$) that is always valid with probability
$1-\delta$; and the width of the interval converges to zero at a
$\sqrt{n}$ rate, where $n$ is the length of the sample path.

%We provide confidence guarantees that are fully data-dependent in the sense that they yield non-trivial
%finite-sample deviation bounds without any knowledge of the mixing coefficients.
%
%
%The novelty of our approach lies in several key aspects.
%Crucially,
%we do not assume any a priori bounds
%on the mixing time,
%which
%sets our approach apart from
%most others that do
%require some advance knowledge of this quantity.
%Another
%advantage of our result over
%common practice is the sufficiency of a single long sampling sequence,
%without access to a restart mechanism.
%Perhaps most importantly, we overcome an obstacle inherent in the chicken-and-egg nature of
%estimating mixing coefficients: to give confidence bounds one typically needs to know the mixing time.
%We provide confidence guarantees that are fully data-dependent in the sense that they yield non-trivial
%finite-sample deviation bounds without any knowledge of the mixing coefficients.
%%Upper and lower sample complexity bounds are provided and challenging open problems are posed.
%%\todod{I would remove the last sentence, at least the part about
%%``challenging'' open problems.}
%%%test