### 2015-06-02 matrix-est.tex

%\documentclass[11pt]{article}
%\documentclass[anon,12pt]{colt2015} 
\documentclass{article}
\usepackage{nips15submit_e,times}
\input{__includes}
\renewcommand\citet\cite % since NIPS is using numbered refs
\renewcommand\citep\cite % since NIPS is using numbered refs

%\title{Single sample path based estimation of the mixing time in finite, reversible Markov chains}
\title{%
  Mixing time estimation in reversible Markov chains from a single
  sample path%
}

\author{
David S.~Hippocampus\thanks{ Use footnote for providing further information
about author (webpage, alternative address)---\emph{not} for acknowledging
funding agencies.} \\
Department of Computer Science\\
Cranberry-Lemon University\\
Pittsburgh, PA 15213 \\
\texttt{hippo@cs.cranberry-lemon.edu} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\AND
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
(if needed)\\
}

\begin{document}
\begin{bibunit}[myunsrt]
  \maketitle

  \begin{abstract} 
  \input{abstract}
  \end{abstract}

  \if0
  \begin{keywords}
    Markov chain; mixing time; spectral gap;
    fully data-dependent confidence interval
  \end{keywords}
  \fi

  \section{Introduction}\label{sec:intro}
  \input{intro-nips}

  \section{Notation}\label{sec:notation}
  \input{notation}

  \section{Setting}\label{sec:setting}
  \input{setting}

  \section{Point estimation}\label{sec:rates}
  \input{rates}

  \section{Fully empirical confidence intervals}\label{sec:empirical}
  \input{empirical}

  \section{Discussion}\label{sec:discussion}
  \input{conclusions}

  \subsubsection*{References}
  {\def\section*#1{}\small\putbib[all]}
\end{bibunit}

\newpage

\appendix

\begin{bibunit}[myunsrt]
  \section{Proof of \cref{thm:err}}\label{app:upper}
  \input{upperbound-new}

  \section{Proofs of the lower bounds}\label{app:lower}
  \input{lowerbound}

  \section{Proof of \cref{thm:empirical}}\label{app:empirical}
  \input{appendix-empirical}

  \section{Proof of \cref{thm:combined}}\label{app:combined}
  \input{appendix-combined}

  \subsubsection*{References}
  {\def\section*#1{}\small\putbib[all]}
\end{bibunit}

\end{document}