### 2015-06-09 arxiv-v1/appendix-combined.tex
arxiv version?

\newcommand\LB{\ensuremath{\operatorname{lb}}}
\newcommand\piminlb{\ensuremath{\hat\pi_{\star,\LB}}}
\newcommand\gaplb{\ensuremath{\hat\gamma_{\star,\LB}}}
Let $\piminlb$ and $\gaplb$ be the lower bounds on $\pimin$ and
$\gap$, respectively, computed from \cref{alg:empest}.
Let $\hatpimin$ and $\hatgap$ be the estimates of $\pimin$ and $\gap$
computed using the estimators from \cref{thm:err}.
By a union bound, we have by \cref{thm:err} and \cref{thm:empirical}
that with probability at least $1-2\delta$,
\begin{equation}
  \Abs{\hatpimin-\pimin}
  \le
  C \,
  \Parens{
    \sqrt{\frac{\pimin\log\frac{d}{\piminlb\delta}}{\gaplb n}}
    +
    \frac{\log\frac{d}{\piminlb\delta}}{\gaplb n}
  }
  \label{eq:pimin-plugin}
\end{equation}
and
\begin{equation}
  \Abs{\hatgap-\gap}
  \leq
  C \,
  \Parens{
    \sqrt{\frac{\log\frac{d}{\delta}\cdot\log\frac{n}{\piminlb\delta}}{\piminlb\gaplb n}}
    +
    \frac{\log\frac{d}{\delta}\cdot\log\frac{n}{\piminlb\delta}}{\piminlb\gaplb n}  
    + \frac{\log\frac{1}{\gaplb}}{\gaplb n}  
  }
  \,.
  \label{eq:gap-plugin}
\end{equation}
The bound on $\abs{\hatgap-\gap}$ in \cref{eq:gap-plugin}---call it
$\hat{w}'$---is fully observable and hence yields a confidence
interval for $\gap$.
The bound on $\abs{\hatpimin-\pimin}$ in \cref{eq:pimin-plugin}
depends on $\pimin$, but from it one can derive
\[
  \Abs{\hatpimin-\pimin}
  \le
  C' \,
  \Parens{
    \sqrt{\frac{\hatpimin\log\frac{d}{\piminlb\delta}}{\gaplb n}}
    +
    \frac{\log\frac{d}{\piminlb\delta}}{\gaplb n}
  }
\]
using the approach from the proof of \cref{lem:P-obs-bound}.
Here, $C'>0$ is an absolute constant that depends only on $C$.
This bound---call it $\hat{b}'$---is now also fully observable.
We have established that in the $1-2\delta$ probability event from
above,
\[
  \pimin \in \wh{U} := [\hatpimin-\hat{b}', \hatpimin+\hat{b}']
  , \quad
  \gap \in \wh{V} := [\hatgap-\hat{w}', \hatgap+\hat{w}']
  .
\]
It is easy to see that almost surely (as $n \to \infty$),
\[
  \sqrt{n / \log(n)} \hat{w}'
  = O\Parens{
    \sqrt{\frac{\log(d/\delta)}{\pimin\gap}}
  }
\]
and
\[
  \sqrt{n} \hat{b}'
  = O\Parens{
    \sqrt{\frac{\pimin \log\frac{d}{\pimin\delta}}{\gap}}
  }
  .
\]
This completes the proof of \cref{thm:combined}.
\hfill\qed