### 2015-06-09 arxiv-v1/paper.tex

\documentclass{article}
\input{__includes}
\renewcommand\citet\cite % since NIPS is using numbered refs
\renewcommand\citep\cite % since NIPS is using numbered refs

\title{%
  Mixing Time Estimation in Reversible Markov Chains from a Single
  Sample Path%
}

\author[1]{\mbox{Daniel Hsu}}
\author[2]{\mbox{Aryeh Kontorovich}}
\author[3]{\mbox{Csaba Szepesv\'ari}}
\affil[1]{\mbox{Columbia University}}
\affil[2]{\mbox{Ben-Gurion University}}
\affil[3]{\mbox{University of Alberta}}

\begin{document}
\maketitle

\begin{abstract} 
\input{abstract}
\end{abstract}

\section{Introduction}\label{sec:intro}
\input{intro}

\section{Preliminaries}\label{sec:prelim}
\input{setting}

\section{Point estimation}\label{sec:rates}
\input{rates}

\section{Fully empirical confidence intervals}\label{sec:empirical}
\input{empirical}

\section{Discussion}\label{sec:discussion}
\input{conclusions}

\bibliography{all}
\bibliographystyle{plain}

\appendix

\section{Proof of \cref{thm:err}}\label{app:upper}
\input{upperbound}

\section{Proofs of the lower bounds}\label{app:lower}
\input{lowerbound}

\section{Proof of \cref{thm:empirical}}\label{app:empirical}
\input{appendix-empirical}

\section{Proof of \cref{thm:combined}}\label{app:combined}
\input{appendix-combined}

\end{document}