### 2015-06-09 arxiv-v1/paper.tex
author header

\documentclass{article}
\input{__includes}
\renewcommand\citet\cite
\renewcommand\citep\cite

\title{%
  Mixing Time Estimation in Reversible Markov Chains from a Single
  Sample Path%
}

\author{%
  Daniel Hsu \\
  Columbia University \\
  \url{djhsu@cs.columbia.edu}
  \and
  Aryeh Kontorovich \\
  Ben-Gurion University \\
  \url{karyeh@cs.bgu.ac.il}
  \and
  Csaba Szepesv\'ari \\
  University of Alberta \\
  \url{szepesva@cs.ualberta.ca}
}

\begin{document}
\maketitle

\begin{abstract} 
\input{abstract}
\end{abstract}

\section{Introduction}\label{sec:intro}
\input{intro}

\section{Preliminaries}\label{sec:prelim}
\input{setting}

\section{Point estimation}\label{sec:rates}
\input{rates}

\section{Fully empirical confidence intervals}\label{sec:empirical}
\input{empirical}

\section{Discussion}\label{sec:discussion}
\input{conclusions}

\bibliography{all}
\bibliographystyle{plain}

\appendix

\section{Proof of \cref{thm:err}}\label{app:upper}
\input{upperbound}

\section{Proofs of the lower bounds}\label{app:lower}
\input{lowerbound}

\section{Proof of \cref{thm:empirical}}\label{app:empirical}
\input{appendix-empirical}

\section{Proof of \cref{thm:combined}}\label{app:combined}
\input{appendix-combined}

\end{document}