### 2015-08-03 nips2015/rebuttal.txt
prelim rebuttal

We thank the referees for the helpful comments. A detailed response follows.

Tightness of bounds: We don't claim tightness of the bounds -- they
are indeed rather loose -- but our results give the first non-trivial
(upper and lower) bounds in this challenging setting. Our work are
motivated by practical applications where fully data-dependent
intervals were sought but not actually achieved. Our results are a
first step towards a practical solution: we demonstrate feasibility
and also prove limitations on what is possible.

Relaxation vs mixing time: In the revision, we will be more explicit
about estimating the relaxation time rather than mixing time proper
(throughout the abstract and paper). We agree that this is an
important distinction that should be emphasized.

Reviewer_1

3. Qualitatively, Theorems 1 and 2 show that if n is less than some
function of (d,pi,gamma), then for any estimator based on a sample
path of length n, there is some Markov chain such that the error is
large.

4. If n > max{16C^2,4C}/(pi_* gamma_*), then the RHS in (4) (without
the logarithmic factor) is less than pi_*/2.

If n > max{16C^2,4C}/(pi_* gamma_*^3), then the RHS in (5) (without
the logarithmic factor) is less than gamma_*/2.

5. Combining Theorems 1 and 2 with a standard argument yields the
lower bound in the introduction. We will include the details in the
revision.

6.

7. To conclude Theorem 2, it follows from Paley-Zygmund and the moment
bounds that, with probability > 1/4, the earliest time in which a
sample path visits all d states is at least c d log(d) / \bar\gamma.
This implies that an estimator using a shorter sample path cannot
distinguish P^0 from P^i for some i (with probability at least 1/4),
and hence must have large estimation error.

Reviewer_2

\tilde{O}: We'll use this more standard use in the revision.

Theorem 2: Indeed, including an extra constant factor in the claim (or
minding inequality strictness) resolves this issue.

Lemma 2: We'll include more details in the revision.

Theorem 2 (line 958): The estimator cannot rule out P^i without having
vistited state i. We'll clarify this point in the revision.

Reviewer_7

Perfect sampling: Indeed, this line of work is related, although these
methods generally do not apply in our setting of estimation from a
single sample path. We'll discuss this in the revision.