### 2015-08-05 nips2015/rebuttal.txt
rebuttal

We thank the referees for the helpful comments. We'll incorporate the
suggestions in a revision of the paper, especially where more detailed
explanations are needed.

The following two concerns were common to many of the reviews.

Tightness of bounds: We don't claim tightness of the bounds -- they
are indeed rather loose -- but our results give the first non-trivial
(upper and lower) bounds in this challenging setting. Our work are
motivated by practical applications where fully data-dependent
intervals were sought but not actually achieved. Our results are a
first step towards a practical solution: we demonstrate feasibility
and also prove limitations on what is possible.

Relaxation vs mixing time: In the revision, we will be more explicit
about estimating the relaxation time rather than mixing time proper
(throughout the abstract and paper). We fully agree that this is a
very important distinction that should be emphasized, and we'll amend
the overstated claims throughout.

Additional responses to individual reviews follow.

Reviewer_1

3. Qualitatively, Theorems 1 and 2 show that if n is less than some
function of (d,pi,gamma), then for any estimator based on a sample
path of length n, there is some Markov chain such that the error is
large.

4. If n > max{16C^2,4C}/(pi_* gamma_*), then the RHS in (4) (without
the logarithmic factor) is less than pi_*/2.

If n > max{16C^2,4C}/(pi_* gamma_*^3), then the RHS in (5) (without
the logarithmic factor) is less than gamma_*/2.

We will add these details in the revision.

5. Combining Theorems 1 and 2 with a standard argument yields the
lower bound in the introduction. We will include the details in the
revision.

6. ???

7. To conclude Theorem 2, it follows the second-moment method (e.g.,
Paley-Zygmund) and the moment bounds that, with probability > 1/4, the
earliest time in which a sample path visits all d states is at least c
d log(d) / \bar\gamma.  This implies that an estimator using a shorter
sample path cannot distinguish P^0 from P^i for some i (with
probability > 1/4), and hence must have large estimation error. We'll
include these details in the revision.

Reviewer_2

\tilde{O}: We'll adopt this more standard use in the revision.

Theorem 2: Indeed, including an extra constant factor in the claim (or
minding inequality strictness) resolves this issue.

Theorem 2 (line 958): The estimator cannot rule out P^i without having
vistited state i. We'll clarify this point in the revision.

Reviewer_7

Perfect sampling: Indeed, this line of work is related, although these
methods generally do not apply in our setting of estimation from a
single sample path. We'll discuss this in the revision.