### 2015-08-07 nips2015/rebuttal.txt
.

We thank the referees for the helpful comments. We'll incorporate the
suggestions in a revision of the paper, especially where more detailed
explanations are needed.

The following two concerns were common to multiple reviews.

Tightness of bounds: We don't claim tightness of the bounds -- they
are indeed rather loose -- but our results give the first non-trivial
(upper and lower) bounds in this challenging setting. Our work is
motivated by practical applications where fully data-dependent
intervals were sought but not actually achieved. Our results are a
first step towards a practical solution: we demonstrate feasibility
and also prove limitations on what is possible.

Relaxation vs mixing time: In the revision, we will be more explicit
about estimating the relaxation time rather than mixing time proper
(throughout the abstract and paper). We fully agree that this is a
very important distinction that should be emphasized, and we'll amend
the overstated claims throughout.

Additional responses to individual reviews follow.

Reviewer_1

2. We'll make this edit in the revision.

3. Qualitatively, Theorems 1 and 2 show that if n is less than some
function of (d,pi,gamma), then for any estimator based on a sample
path of length n, there is some Markov chain such that the error is
large. We'll add additional clarifying interpretations in the
revision.

4. If n > max{16C^2,4C}/(pi_* gamma_*), then the RHS in (4) (without
the logarithmic factor) is less than pi_*/2.

If n > max{16C^2,4C}/(pi_* gamma_*^3), then the RHS in (5) (without
the logarithmic factor) is less than gamma_*/2.

We will add these details in the revision.

5. Combining Theorems 1 and 2 with a standard argument yields the
lower bound in the introduction. We will include the details in the
revision.

6. Unfortunately, we were not able to understand the counterexample,
likely due to typos in the review. We hope to resolve this via the
program chairs/area chair.

7. To conclude Theorem 2, it follows the second-moment method (e.g.,
Paley-Zygmund) and the moment bounds that, with probability > 1/4, the
earliest time in which a sample path visits all d states is at least c
d log(d) / \bar\gamma.  This implies that an estimator using a shorter
sample path cannot distinguish P^0 from P^i for some i (with
probability > 1/4), and hence must have large estimation error. We'll
include these details in the revision.

Reviewer_2

Theorem 5: We get an improvement when the min in the bound for
|\hat{V}| is achieved by the first term. A sheepish caveat: this is
not always the case for all sufficiently large n, due to the
sqrt{log(n)} term in the numerator, but will be true for a very large
range of n. There is more discussion in Section 5.

\tilde{O}: We'll adopt this more standard use in the revision.

kappa: It is defined just above Theorem 4.

Theorem 2: Indeed, including an extra constant factor in the claim (or
minding inequality strictness) resolves this issue.

Theorem 2 (line 958): An estimator cannot rule out P^i without having
vistited state i. We'll clarify this point in the revision.

Reviewer_7

Perfect sampling: Indeed, this line of work is related, although these
methods generally do not apply in our setting of estimation from a
single sample path. We'll discuss this in the revision.