### 2015-08-09 nips2015/rebuttal.txt
Minor changes to rebuttal

We thank the referees for the helpful comments. We'll incorporate the suggestions in a revision of the paper, especially where more detailed explanations are needed.

The following two concerns were common to multiple reviews.

Tightness of bounds: We don't claim tightness of the bounds (we did not attempt to optimize the constants) but our results give the first non-trivial (upper and lower) bounds in this challenging problem. In particular, the estimator is based on a less than straighforward, indirect approach while avoiding slow rates (as explained at the beginning of Section 4) that builds on advanced perturbation results from the theory of stochastic matrices. Our work is motivated by practical applications where fully data-dependent intervals were sought but not actually achieved. Our results are a first step towards a practical solution: we demonstrate feasibility and also prove limitations on what is possible, both add significantly to our knowledge.

Relaxation vs mixing time: In the revision, we will be more explicit about estimating the relaxation time rather than mixing time proper (throughout the abstract and paper). We fully agree that this is a very important distinction that should be emphasized, and we'll amend the overstated claims throughout.

Additional responses to individual reviews follow.

Reviewer_1

2. We'll make this edit in the revision.

3. Qualitatively, Theorems 1 and 2 show that if n is less than some function of (d,pi,gamma), then for any estimator based on a sample path of length n, there is some Markov chain such that the error is large. We'll add additional clarifying interpretations in the revision.

4. If n > max{16C^2,4C}/(pi_* gamma_*), then the RHS in (4) (without the logarithmic factor) is less than pi_*/2.

If n > max{16C^2,4C}/(pi_* gamma_*^3), then the RHS in (5) (without the logarithmic factor) is less than gamma_*/2.

We will add these details in the revision.

5. Combining Theorems 1 and 2 with a standard argument yields the lower bound in the introduction. We will include the details in the revision.

6. Regarding Theorem 1: Note that "c" is an absolute constant in the theorem, not something that can be chosen arbitrarily. In particular, one value that works in the theorem is c=1/4.

There were some minor numerical calculation errors in the proof of Theorem 1, which we can clarify here. We need to change the "1/6" in the theorem statement and proof to "1/8". Furthermore, in the proof, \bar\pi should be taken in (0,1/4), and we should set \varepsilon := \bar\pi. Observe that \pi^{(2)} = (1/(1+2\varepsilon),2\varepsilon/(1+2\varepsilon)). Thus, \pi_\star \geq \bar\pi in both P^{(1)} and P^{(2)}. To guarantee |\hat\gamma_\star - \gamma_\star| < 1/8, it must be possible to distinguish the two Markov chains. The rest of the proof is correct as is.

7. The conclusion of Theorem 2 follows using the second-moment method (e.g., Paley-Zygmund) and the moment bounds that, with probability > 1/4, the earliest time in which a sample path visits all d states is at least c d log(d) / \bar\gamma.  This implies that an estimator using a shorter sample path cannot distinguish P^0 from P^i for some i (with probability > 1/4), and hence must have large estimation error. We'll include these details in the revision.

Minor comment 3: We respectfully disagree. The definition of a confidence interval does not distinguish intervals based on what is known, which itself may depend on the application (in specific applications additional knowledge if often available).

Reviewer_2

Theorem 5: We get an improvement when the min in the bound for |\hat{V}| is achieved by the first term. A sheepish caveat: this is not always the case for all sufficiently large n, due to the sqrt{log(n)} term in the numerator, but will be true for a very large range of n. There is more discussion in Section 5.

Theorem 2 (line 188): Indeed, including an extra constant factor in the claim resolves this issue.

\tilde{O}: We'll adopt this more standard use in the revision.

kappa: It is defined just above Theorem 4.

line 926: Thanks. See response to Issue 6 by Reviewer 1.

Theorem 2 (line 958): An estimator cannot rule out P^i without having visited state i. We'll clarify this point in the revision.

Reviewer_7

Perfect sampling: Indeed, this line of work is related, although these methods generally do not apply in our setting of estimation from a single sample path. We'll discuss this in the revision.