### 2015-10-15 nips2015ws/rates.tex

%We present lower/upper bounds on rates
%for estimating spectral gap as a function of the length of the
%sample path $n$.

\subsection{Lower bounds}
\label{sec:rates-lower}
We first show lower bounds on the number of observations necessary to
achieve a fixed multiplicative (or even just additive) accuracy in
estimating the spectral gap $\gap$.
By \cref{eq:mixing-time-bound}, the multiplicative accuracy lower
bound for $\gap$ gives the same lower bound for estimating the mixing
time.
Our first result holds even for two state Markov chains and shows that a sequence length of $\Omega(1/\pimin)$
is necessary to achieve even a constant \emph{additive} accuracy in estimating $\gap$.
\begin{theorem}
  \label{thm:lb-pimin}
  Pick any $\bar\pi \in (0,1/4)$.
  Consider any estimator $\hatgap$ that takes as input a random sample
  path of length $n \leq 1/(4\bar\pi)$ from a Markov chain starting
  from any desired initial state distribution.
  There exists a two-state ergodic and reversible Markov chain
  distribution with spectral gap $\gap \geq 1/2$ and minimum
  stationary probability $\pimin \geq \bar\pi$ such that
  $\Pr\Brackets{ |\hatgap - \gap| \geq 1/8 } \geq 3/8$.
%  \[
%    \Pr\Brackets{ |\hatgap - \gap| \geq 1/8 } \geq 3/8 .
%  \]
\end{theorem}
Next, considering $d$ state chains, we show that 
a sequence of length $\Omega(d\log(d)/\gap)$ is required
to estimate $\gap$ up to a constant multiplicative accuracy.
Essentially, the sequence may have to visit all $d$ states at least
$\log(d)/\gap$ times each, on average.
This holds \emph{even} if $\pimin$ is within a factor of two of the
\emph{largest} possible value of $1/d$ that it can take, i.e., when
$\vpi$ is nearly uniform.
\begin{theorem}
  \label{thm:lb-gap}
  There is an absolute constant $c>0$ such that the following holds.
  Pick any positive integer $d \geq 3$ and any $\bar\gamma \in
  (0,1/2)$.
  Consider any estimator $\hatgap$ that takes as input a random sample
  path of length $n < c d\log(d) / \bar\gamma$ from a $d$-state
  reversible Markov chain starting from any desired initial state
  distribution.
  There is an ergodic and reversible Markov chain distribution
  with spectral gap $\gap \in [\bar\gamma,2\bar\gamma]$ and minimum
  stationary probability $\pimin \geq 1/(2d)$ such that
  $\Pr\Brackets{ |\hatgap - \gap| \geq \bar\gamma/2} \geq 1/4$.
%  \[
%    \Pr\Brackets{ |\hatgap - \gap| \geq \bar\gamma/2} \geq 1/4 .
%  \]
\end{theorem}

%The proofs of \cref{thm:lb-pimin,thm:lb-gap} are given in
%\cref{app:lower}.

\subsection{A plug-in based point estimator and its accuracy}
\label{sec:rates-upper}
We now consider the problem of estimating $\gap$ using a natural
plug-in estimator.
%For this, we construct a natural plug-in estimator.
%Along the way, we also provide an estimator for the minimum stationary
%probability, allowing one to use the bounds from
%\cref{eq:mixing-time-bound} to trap the mixing time.

\iffalse
Define the random matrix $\wh\vM \in [0,1]^{d \times d}$ and random
vector $\hat\vpi \in \Delta^{d-1}$ by
\begin{align*}
  \wh{M}_{i,j}
  & := \frac{|\{ t \in [n-1] : (X_t,X_{t+1}) = (i,j) \}|}{n-1}
  , \quad i,j \in [d]\,,
  \\
  \hat{\pi}_i
  & := \frac{|\{ t \in [n] : X_t = i \}|}{n}
  , \quad i \in [d]
  \,.
\end{align*}
Furthermore, define
\[
  \Sym(\wh\vL) := \frac12 \parens{ \wh\vL + \wh\vL^\t }
\]
to be the symmetrized version of the (possibly non-symmetric) matrix
\[
  \wh\vL := \Diag(\hat\vpi)^{-1/2} \wh\vM \Diag(\hat\vpi)^{-1/2}
  .
\]
Let $\hat\lambda_1 \geq \hat\lambda_2 \geq \dotsb \geq \hat\lambda_d$
be the eigenvalues of $\Sym(\wh\vL)$.
Our estimator of the minimum stationary probability $\pimin$ is
\[
  \hatpimin := \min_{i \in [d]} \hat\pi_i ,
\]
and our estimator of the spectral gap $\gap$ is
\[
  \hatgap := 1 - \max\{ \hat\lambda_2, |\hat\lambda_d| \} .
\]
These estimators have the following accuracy guarantees:
\fi
\begin{theorem}
  \label{thm:err}
  There exists an absolute constant $C>0$ such that the following
  holds.
  Assume the estimators $\hatpimin$ and $\hatgap$ described above are
  formed from a sample path of length $n$ from an ergodic and
  reversible Markov chain.
  Let $\gap>0$ denote the spectral gap and $\pimin>0$ the minimum
  stationary probability.
  For any $\delta \in (0,1)$, with probability at least $1-\delta$,
  the plug-in estimates $\hatpimin$ and $\hatgap$ of $\pimin$ and
  $\gap$ satisfy
   $\Abs{\hatpimin-\pimin}
    \le
    C \,
    \parens{
      \sqrt{\pimin\log\frac{d}{\pimin\delta}/(\gap n)}
      +
      \log\frac{d}{\pimin\delta}/(\gap n)
    }$
%  \begin{equation}\label{eq:piminbound}
%    \Abs{\hatpimin-\pimin}
%    \le
%    C \,
%    \Parens{
%      \sqrt{\frac{\pimin\log\frac{d}{\pimin\delta}}{\gap n}}
%      +
%      \frac{\log\frac{d}{\pimin\delta}}{\gap n}
%    }
%  \end{equation}
  and
   $\Abs{\hatgap-\gap}
    \leq
    C \,
    \parens{
      \sqrt{\log\frac{d}{\delta}\cdot\log\frac{n}{\pimin\delta}/(\pimin\gap n)}
      + \log\frac{1}{\gap}/(\gap n)
    }$.
%  \begin{equation}\label{eq:gapbound}
%    \Abs{\hatgap-\gap}
%    \leq
%    C \,
%    \Parens{
%      \sqrt{\frac{\log\frac{d}{\delta}\cdot\log\frac{n}{\pimin\delta}}{\pimin\gap n}}
%      + \frac{\log\frac{1}{\gap}}{\gap n}  
%    }
%    \,.
%  \end{equation}
\end{theorem}

%\Cref{thm:err} implies that the sequence lengths required to estimate
%$\pimin$ and
%$\gap$ to within constant multiplicative factors are, respectively,
%\[
%  \tilde{O}\Parens{\frac1{\pimin\gap}}
%  \quad\text{and}\quad
%  \tilde{O}\Parens{\frac1{\pimin\gap^3}}
%  .
%\]
%By \cref{eq:mixing-time-bound},
%the second of these is also a bound on the required sequence length to estimate $\tmix$.

%The proof of \cref{thm:err} is based on analyzing the
%convergence of the sample averages $\wh{\vM}$ and
%$\hat\vpi$ to their expectation, and then using perturbation bounds
%for eigenvalues to derive a bound on the error of $\hatgap$.
%However, since these averages are formed using a \emph{single sample
%path} from a (possibly) non-stationary Markov chain, we cannot use
%standard large deviation bounds; moreover applying Chernoff-type
%bounds for Markov chains to each entry of $\wh{\vM}$ would result in a
%significantly worse sequence length requirement, roughly a factor of
%$d$ larger.
%Instead, we adapt probability tail bounds for sums of independent
%random matrices~\citep{tropp2015intro} to our non-iid setting by
%directly applying a blocking technique of~\citet{Bernstein27} as
%described in the article of~\citet{Yu94}.
%Due to ergodicity, the convergence rate can be bounded without any
%dependence on the initial state distribution $\vpi^{(1)}$.
%The proof of \cref{thm:err} is given in \cref{app:upper}.

Note that because the eigenvalues of $\vL$ are the same as that of the
transition probability matrix $\vP$, 
we could have instead opted to
estimate $\vP$, say, using simple frequency estimates obtained from
the sample path, and then computing the second largest eigenvalue of
this empirical estimate $\wh\vP$.
In fact, this approach is a way to extend to non-reversible chains, as
we would no longer rely on the symmetry of $\vM$ or $\vL$.
The difficulty with this approach is that $\vP$ lacks the structure
required by certain strong eigenvalue perturbation results.
One could instead invoke the Ostrowski-Elsner theorem
\citep[cf.~Theorem 1.4 on Page 170 of][]{stewart1990matrix}, which
bounds the \emph{matching distance} between the eigenvalues of a
matrix $\vA$ and its perturbation $\vA+\vE$ by $O(\norm{\vE}^{1/d})$.
Since $\norm{\wh\vP-\vP}$ is expected to be of size $O(n^{-1/2})$,
this approach will give a confidence interval for $\gap$ whose width
shrinks at a rate of $O(n^{-1/(2d)})$---an exponential slow-down
compared to the rate from \cref{thm:err}.
As demonstrated through an example from \citet{stewart1990matrix}, the
dependence on the $d$-th root of the norm of the perturbation cannot
be avoided in general.
Our approach based on estimating a symmetric matrix affords us the use
of perturbation results that exploit more structure.

Returning to the question of obtaining a fully empirical confidence 
interval for $\gap$ and $\pimin$, we notice that,
unfortunately, \cref{thm:err} falls short of being directly suitable for this,
at least without further assumptions.
This is because the deviation terms themselves depend inversely both
on $\gap$ and $\pimin$, and hence can never rule out $0$ (or an
arbitrarily small positive value) as a possibility for $\gap$ or
$\pimin$.\footnote{%
  Using \cref{thm:err}, it is possible to trap $\gap$ in the
  union of \emph{two} empirical confidence intervals---one around
  $\hatgap$ and the other around zero, both of which shrink in width
  as the sequence length increases.%
}
In effect, the fact that the Markov chain could be slow mixing and the
long-term frequency of some states could be small makes it difficult
to be confident in the estimates provided by $\hatgap$ and
$\hatpimin$.
This suggests that in order to obtain fully empirical confidence
intervals, we need an estimator that is not subject to such
effects---we pursue this in \cref{sec:empirical}.
\Cref{thm:err} thus primarily serves as a point of comparison
for what is achievable in terms of estimation accuracy when one does
not need to provide empirical confidence bounds.